Brownian motion of main-belt asteroids on human timescales

### 2025-08-20 response.txt

Response to first referee report.

I thank the referee for their concise and insightful review of the paper.  The
main issue raised is that this paper relies heavily on the companion paper
analysing LSST constraints on asteroid masses from mutual encounters.  That
paper was submitted to PSJ some weeks before this one, and has since been
refereed, revised, accepted, and published in the journal, which I hope will
assuage this worry.

One of the revisions made to that paper addressed an issue also raised by this
referee, which was the rather high value of albedo assumed for the main-belt
asteroid population.  To address this concern for the other paper, we changed
our conversion from (well-observed) H values to (physically relevant) mass
substantially,  Instead of choosing a representative value of  both density and
albedo of the population, we surveyed the literature for individual main-belt
asteroids (MBAs) with mass determinations of any useful precision, then examined
the ratio $M/ 10^{-0.6H}$ directly.  The number of such bodies is actually
larger than the number for which both density and albedo have been determined
to similar precision, so is a simpler and slightly more accurate means of deriving
the "mass-to-light" conversion.  See Figure 1 and Section 2.1 of the published
paper on mass estimation for details.

The adopted result of plotting the objects with known H and M was that the
distribution of mass-to-light ratios of this (incomplete, heterogeneous, almost
certainly biased) population could be crudely characterized by taking a "high"
and "low" value, which differ by a factor of 5.  The main change made to this
Brownian-motion paper is to adopt these two values and recalculate the resulting
integrals that yield the estimated RMS 10-year uncertainties that are the key
result of this paper.  These mass-to-light ratio changes appear as
multiplicative scalings of the previous result.  The thresholds for
individually-detectable impulses in Section 3.2.3 have also been updated in
concert with the updates to the mass-determination paper.

Thus the numerical results presented in this paper have been adjusted, and this
has led to substantial revisions of the sections conveying results, the
discussion sections, and the abstract.  We have changed the main results in
Figure 3 to show both the high- and low-mass MBA cases, and added Table 1 to
summarize the numerical results because there are now a multiplicity of cases to
keep track of.  One important change to the conclusions is that this Brownian
noise is important to consider by people analyzing upcoming Gaia astrometry on
MBAs.

Changed text is in red in the draft, and figures that have changed have this
noted in their captions. As per the referee's comments, the methodology remains
largely unchanged.  Throughout the text, we have made small changes to the
wording of sections where we now see clarity can be improved, and have fixed
some small typos and errors.

@@

A few minor replies to the referee's comments are below.

> This report evaluates the manuscript "Brownian Motion of Main-Belt
> Asteroids on Human Timescales," which examines how gravitational
> interactions between asteroids introduce stochastic positional
> uncertainties over decadal timescales. The study provides a novel
> approach to understanding perturbative effects in asteroid
> ephemerides, with implications for high-precision astronomy.
> 
> The central hypothesis of the study is that the long-term orbital
> evolution of main-belt asteroids is affected by random gravitational
> impulses from other asteroids, leading to a Brownian motion-like
> positional drift. This effect is expected to be particularly
> significant over human timescales (10-100 years), impacting asteroid
> tracking and planetary defense efforts. The study proposes that this
> motion is primarily azimuthal and that current ephemeris models
> underestimate this stochastic contribution.

We did not intend to give the impression that Brownian motion is larger or more
important on decadal scales than over Myr or Gyr scales.  Rather our purpose is to
quantify the effect of the Brownian motion on "real-time" dynamics of the solar
system, i.e. the development of ephemerides that fit obervations and adjust
parameters of the solar system's members and of the laws of gravitation.
Understanding the long-term dynamical consequences of the Brownian motion for
population evolution is beyond our scope.

> 
> The study is based on several key assumptions:
> 
> - Asteroid Encounters as Stochastic Events: The model assumes that
> gravitational encounters between asteroids can be treated as a series
> of independent random impulses.
> 
> - Negligibility of Non-Gravitational Effects: The study assumes that
> radiation forces (e.g., Yarkovsky effect) and thermal recoil forces
> have a minor influence compared to gravitational interactions for
> asteroids larger than ~100 m.
>

Note in section 6.2, we calculate that Yarkovsky effects are significant in
comparison to the Brownian noise for asteroids of diameter >10 km.

> - Mass Uncertainty from LSST Data: The study uses projected LSST data
> to estimate mass uncertainties, heavily relying on Bernstein et
> al. (2025) for these calculations.
> 
> - Choice of Geometric Albedo: The author chooses the geometric albedo
> value that might, in this evaluator's judgment, be too "bright" to be
> representative of the entire population of asteroids in the main
> belt. The problem with that choice lies in the fact that, with that
> value, the mass distribution estimate would be misrepresented. A value
> of 0.15 might be more representative of the average, so I recommend
> its use.
>
